Random number generation and applications on GPU’s in R?

### 2021-10-04 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
no big changes,just to save
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2888 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\subsection{Creating streams}
In \proglang{clRNG} library, a stream object is a structure that contains three states (see the first column in the \proglang{R} output below): the current state of the stream, the initial state (or seed) of the stream , and the initial state of the current substream. The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host computer using the MRG31k3p RNG. If we want to use the streams in work items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work item picks streams from the global memory and copies their current states only to its private memory. In \pkg{gpuRandom}, to avoid unnecessary data transfer between host and device, we are able to create streams both on the host computer and on the GPU device. \fct{gpuRandom::CreateStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on the host as \proglang{R} matrices, so that users can easily view the streams. \fct{gpuRandom::CreateStreamsGpu} is adapted from \fct{clrngMrg31k3pCreateStreams} with the advantage of creating streams directly on GPU device. From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{gpuRandom::CreateStreamsGpu} is more efficient and more frequently used. In the following, we demonstrate the usage of these two functions by generating 4 streams on the host and on the device respectively.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# Create streams on host
streams_on_CPU<-gpuRandom::CreateStreamsCpu(Nstreams=4)
t(streams_on_CPU)
##                 [,1]       [,2]       [,3]       [,4]
## current.g1.1   12345  336690377  502033783  739421137
## current.g1.2   12345  597094797 1322587635 1475938232
## current.g1.3   12345 1245771585 1964121530  730262207
## current.g2.1   12345   85196284 1949818481 1630192198
## current.g2.2   12345  523477687 1607232546  324551134
## current.g2.3   12345 2094976052 1462898381  795289868
## initial.g1.1   12345  336690377  502033783  739421137
## initial.g1.2   12345  597094797 1322587635 1475938232
## initial.g1.3   12345 1245771585 1964121530  730262207
## initial.g2.1   12345   85196284 1949818481 1630192198
## initial.g2.2   12345  523477687 1607232546  324551134
## initial.g2.3   12345 2094976052 1462898381  795289868
## substream.g1.1 12345  336690377  502033783  739421137
## substream.g1.2 12345  597094797 1322587635 1475938232
## substream.g1.3 12345 1245771585 1964121530  730262207
## substream.g2.1 12345   85196284 1949818481 1630192198
## substream.g2.2 12345  523477687 1607232546  324551134
## substream.g2.3 12345 2094976052 1462898381  795289868
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{gpuRandom::CreateStreamsCpu} has only one argument \code{Nstreams}, which is the number of streams to create. For the MRG31k3p RNG, each state of the stream object is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) is by default ``12345''. By setting a seed same as the MRG31k3p default seed in \fct{gpuRandom::CreateStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# Create streams on GPU
seed <- c(12345, 12345, 12345, 12345, 12345, 12345)
streams1 <- gpuRandom::CreateStreamsGpu(seedR=seed, Nstreams=4, keepInitial=TRUE)
streams_from_GPU <- as.matrix(streams1)
streams2 = deepcopy(streams1)
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
t(streams_from_GPU)   
##                 [,1]       [,2]       [,3]       [,4]
## current.g1.1   12345  336690377  502033783  739421137
## current.g1.2   12345  597094797 1322587635 1475938232
## current.g1.3   12345 1245771585 1964121530  730262207
## current.g2.1   12345   85196284 1949818481 1630192198
## current.g2.2   12345  523477687 1607232546  324551134
## current.g2.3   12345 2094976052 1462898381  795289868
## initial.g1.1   12345  336690377  502033783  739421137
## initial.g1.2   12345  597094797 1322587635 1475938232
## initial.g1.3   12345 1245771585 1964121530  730262207
## initial.g2.1   12345   85196284 1949818481 1630192198
## initial.g2.2   12345  523477687 1607232546  324551134
## initial.g2.3   12345 2094976052 1462898381  795289868
## substream.g1.1 12345  336690377  502033783  739421137
## substream.g1.2 12345  597094797 1322587635 1475938232
## substream.g1.3 12345 1245771585 1964121530  730262207
## substream.g2.1 12345   85196284 1949818481 1630192198
## substream.g2.2 12345  523477687 1607232546  324551134
## substream.g2.3 12345 2094976052 1462898381  795289868
\end{verbatim}
\end{kframe}
\end{knitrout}
The \fct{gpuRandom::CreateStreamsGpu} has three arguments:
\begin{itemize}
  \item \code{seedR}: an R vector of length 6, which is the seed of streams. Default is set to a vector of six ``12345''.
  \item \code{Nstreams}: number of streams to create.
  \item \code{keepInitial}: logical specifying whether to keep the seed in generated streams. Default is set to TRUE.
\end{itemize}
\fct{CreateStreamsGpu} in \pkg{gpuRandom} are similar to \code{.Random.seed} in \proglang{R}. By setting the RNG status with the \fct{set.seed} function in \proglang{R}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs. Similarly in \pkg{gpuRandom}, with the \code{seedR} argument, we are able to produce exactly the same sequence of streams as many times as we want.

@@

Now we use the 4 streams created in Section 2.1 and an NDRange of 1 by 4 work items to generate a vector \code{sim_1} of 6 i.i.d. $U (0,1)$ random numbers using the function \fct{gpuRandom::runif}. If \code{streams} and \code{Nglobal} are both missing, then by default \code{Nglobal} is set as (64,8), and a `vclMatrix' of \code{prod(Nglobal)} streams is supplied. The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
sim_1 = gpuRandom::runif(n = 6, streams = streams1, Nglobal = c(1, 4),
    type = "double")
as.vector(sim_1)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
\end{verbatim}
\end{kframe}
The arguments of the \fct{gpuRandom::runif} are described as follows:
\begin{itemize}
  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or specifying the length if to create a vector.
  \item \code{streams}: streams used for random number generation.
  \item \code{Nglobal}: global index space that defines the total work-items that execute in parallel. Default is set as $(64,8)$.
  \item \code{type}: ``double'' or ``float'' format of generated random numbers.
\end{itemize}
Next we show that by using a same stream \code{streams2} as \code{streams1},  \fct{gpuRandom::runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
sim_2 <- gpuRandom::runif(n = 6, streams = streams2, Nglobal = c(1, 4),
    type = "double")
as.vector(sim_2)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
\end{verbatim}
\end{kframe}
Notice that \code{streams2} is a \code{deepcopy} of \code{streams1} before \code{streams1} is used. Reusing \code{streams1} will produce a vector different from \code{sim_1} (see \code{sim_3}). Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the first six rows (current states) of \code{streams1} after it being used.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
sim_3 = gpuRandom::runif(n = 6, streams = streams1, Nglobal = c(1, 4),
    type = "double")
as.vector(sim_3)
## [1] 0.6487742 0.1112075 0.3661944 0.5018562 0.1088229 0.3114331
t(as.matrix(streams1))[1:6, ]
##            [,1]       [,2]       [,3]       [,4]
## [1,] 1167281028  640923766  502033783  739421137
## [2,] 1918428443 1912157188 1322587635 1475938232
## [3,] 1858462085  286315187 1964121530  730262207
## [4,]  933585541 2119609883 1949818481 1630192198
## [5,] 1132031887  834429287 1607232546  324551134
## [6,]  465230163   47498872 1462898381  795289868
\end{verbatim}
\end{kframe}
Unlike the objects created by \proglang{R}, by default they remain in the memory after a restart. Streams created on GPU does not remain in the memory when a new \proglang{R} session begins. However, we can save the streams on the CPU in a file using the \fct{save} function, and later recall the streams with the \fct{load} function, in this way we can reproduce the same sequence of random numbers in simulations. Below is a trivial example, in which we demonstrate saving \code{streams1} to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU to produce the vector called \code{sim_saved}, which is also the same as \code{sim_1}.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
saveRDS(streams_from_GPU, "streams_from_GPU.rds")
\end{verbatim}
\end{kframe}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# Load the streams object as streams_saved
streams_saved <- gpuR::vclMatrix(readRDS("streams_from_GPU.rds"))
sim_saved = gpuRandom::runif(n=6,streams=streams_saved, 
                             Nglobal=c(1,4), type="double")
as.vector(sim_saved)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
\end{verbatim}
\end{kframe}

We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}). The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{gpuRandom::runif} does come from a standard uniform distribution.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streams <- CreateStreamsGpu(Nstreams =512, keepInitial=1)
random_vector<-gpuRandom::runif(n=50000, streams=streams, 
                                Nglobal=c(64,8), type="double")
random_vector<-as.vector(random_vector)
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
\includegraphics[width=0.45\textwidth]{figure/largenumberuniforms-1} 
\includegraphics[width=0.45\textwidth]{figure/largenumberuniforms-2} 
\end{knitrout}
\caption{histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
\end{figure}

@@

In the following code, We create a $1024 \times 512$ matrix named \code{normal_matrix} of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}. The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution. We find doing a Q-Q plot can occupy a large amount of time in the process of producing an R markdown file, so as an incidental work, we adapted the \proglang{R} function \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on GPU, which can speed up the Q-Q plot greatly.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streams <- CreateStreamsGpu(Nstreams =512*128, keepInitial=1)
normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
                                Nglobal=c(512,128), type="float")
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
avector<-as.vector(as.matrix(normal_matrix))
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
hist(avector,breaks=40, main = "histogram")
gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
\end{verbatim}
\end{kframe}
\includegraphics[width=0.45\textwidth]{figure/histAndQQ-1} 
\includegraphics[width=0.45\textwidth]{figure/histAndQQ-2} 
\end{knitrout}
\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
\end{figure}

@@

\subsection{Comparing running time: Month data example}

>>>>

\subsection{Creating streams}

{\tt changed createStreams functions to be compatible with each other}

create streams on cpu
\begin{knitrout}\tiny
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
myStreamsCpu <- createStreamsCpu(n=4, seed = 12345)
t(myStreamsCpu)
##               [,1]       [,2]       [,3]       [,4]
## current.g1.1 12345  336690377  502033783  739421137
## current.g1.2 12345  597094797 1322587635 1475938232
## current.g1.3 12345 1245771585 1964121530  730262207
## current.g2.1 12345   85196284 1949818481 1630192198
## current.g2.2 12345  523477687 1607232546  324551134
## current.g2.3 12345 2094976052 1462898381  795289868
## initial.g1.1 12345  336690377  502033783  739421137
## initial.g1.2 12345  597094797 1322587635 1475938232
## initial.g1.3 12345 1245771585 1964121530  730262207
## initial.g2.1 12345   85196284 1949818481 1630192198
## initial.g2.2 12345  523477687 1607232546  324551134
## initial.g2.3 12345 2094976052 1462898381  795289868
\end{verbatim}
\end{kframe}
\end{knitrout}

streams on gpu
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
myStreamsGpu1 = vclMatrix(myStreamsCpu)
myStreamsGpu2 = createStreams(n=4, seed=12345)
range(as.matrix(myStreamsGpu1 - myStreamsGpu2))
## [1] 0 0
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
sim_2 <- gpuRandom::runif(n = 6, streams = myStreamsGpu1, Nglobal = c(1,
    nrow(myStreamsGpu1)), type = "double")
as.vector(sim_2)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
\end{verbatim}
\end{kframe}
\end{knitrout}

check seed
\begin{knitrout}\tiny
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu))
##      current.g1.1 current.g1.2 current.g1.3 current.g2.1 current.g2.2
## [1,]    878672095    240667857    240667857    642281259   1069151070
## [2,]   2113333390    559530223   1309565828   1335994581     61444481
## [3,]    502033783   1322587635   1964121530   1949818481   1607232546
## [4,]    739421137   1475938232    730262207   1630192198    324551134
##      current.g2.3 initial.g1.1 initial.g1.2 initial.g1.3 initial.g2.1
## [1,]    809054265        12345        12345        12345        12345
## [2,]    197003928    336690377    597094797   1245771585     85196284
## [3,]   1462898381    502033783   1322587635   1964121530   1949818481
## [4,]    795289868    739421137   1475938232    730262207   1630192198
##      initial.g2.2 initial.g2.3
## [1,]        12345        12345
## [2,]    523477687   2094976052
## [3,]   1607232546   1462898381
## [4,]    324551134    795289868
\end{verbatim}
\end{kframe}
\end{knitrout}

check reproducibility
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
as.vector(gpuRandom::runif(n = 4, streams = myStreamsGpu1, Nglobal = c(1,
    nrow(myStreamsGpu1)), type = "double"))
## [1] 0.6487742 0.1112075 0.3661944 0.5018562
as.vector(gpuRandom::runif(n = 10, streams = myStreamsGpu2, Nglobal = c(1,
    nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
## [1] 0.6487742 0.1112075 0.3661944 0.5018562
\end{verbatim}
\end{kframe}
\end{knitrout}

In \proglang{clRNG} library, a stream object is a structure that contains three states (see the first column in the \proglang{R} output below): the current state of the stream, the initial state (or seed) of the stream , and the initial state of the current substream. The function \fct{clrngMrg31k3pCreateStreams} from \proglang{clRNG} creates streams on the host computer using the MRG31k3p RNG. If we want to use the streams in work items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work item picks streams from the global memory and copies their current states only to its private memory. In \pkg{gpuRandom}, to avoid unnecessary data transfer between host and device, we are able to create streams both on the host computer and on the GPU device. \fct{gpuRandom::CreateStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on the host as \proglang{R} matrices, so that users can easily view the streams. \fct{gpuRandom::CreateStreamsGpu} is adapted from \fct{clrngMrg31k3pCreateStreams} with the advantage of creating streams directly on GPU device. From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{gpuRandom::CreateStreamsGpu} is more efficient and more frequently used. In the following, we demonstrate the usage of these two functions by generating 4 streams on the host and on the device respectively.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# Create streams on host
streams_on_CPU<-gpuRandom::CreateStreamsCpu(seed = rep(12345, 6), Nstreams=4)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'CreateStreamsCpu' is not an exported object from 'namespace:gpuRandom'}}\begin{verbatim}
t(streams_on_CPU)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 't': object 'streams\_on\_CPU' not found}}\end{kframe}
\end{knitrout}
\fct{gpuRandom::CreateStreamsCpu} has only one argument \code{Nstreams}, which is the number of streams to create. For the MRG31k3p RNG, each state of the stream object is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) is by default ``12345''. By setting a seed same as the MRG31k3p default seed in \fct{gpuRandom::CreateStreamsGpu}, we can produce on device exactly the same streams as those created on host as follows,
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# Create streams on GPU
seed <- c(12345, 12345, 12345, 12345, 12345, 12345)
streams1 <- gpuRandom::CreateStreamsGpu(seedR=seed, Nstreams=4, keepInitial=TRUE)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'CreateStreamsGpu' is not an exported object from 'namespace:gpuRandom'}}\begin{verbatim}
streams_from_GPU <- as.matrix(streams1)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.matrix(streams1): object 'streams1' not found}}\begin{verbatim}
streams2 = deepcopy(streams1)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'object' in selecting a method for function 'deepcopy': object 'streams1' not found}}\end{kframe}
\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gpuR::colnames(streams\_from\_GPU) = c("{}current.g1.1"{}, "{}current.g1.2"{}, : object 'streams\_from\_GPU' not found}}\end{kframe}
\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
t(streams_from_GPU)   
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 't': object 'streams\_from\_GPU' not found}}\end{kframe}
\end{knitrout}
The \fct{gpuRandom::CreateStreamsGpu} has three arguments:
\begin{itemize}
  \item \code{seedR}: an R vector of length 6, which is the seed of streams. Default is set to a vector of six ``12345''.
  \item \code{Nstreams}: number of streams to create.
  \item \code{keepInitial}: logical specifying whether to keep the seed in generated streams. Default is set to TRUE.
\end{itemize}
\fct{CreateStreamsGpu} in \pkg{gpuRandom} are similar to \code{.Random.seed} in \proglang{R}. By setting the RNG status with the \fct{set.seed} function in \proglang{R}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs. Similarly in \pkg{gpuRandom}, with the \code{seedR} argument, we are able to produce exactly the same sequence of streams as many times as we want.

@@

Now we use the 4 streams created in Section 2.1 and an NDRange of 1 by 4 work items to generate a vector \code{sim_1} of 6 i.i.d. $U (0,1)$ random numbers using the function \fct{gpuRandom::runif}. If \code{streams} and \code{Nglobal} are both missing, then by default \code{Nglobal} is set as (64,8), and a `vclMatrix' of \code{prod(Nglobal)} streams is supplied. The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
sim_1 = gpuRandom::runif(n = 6, streams = streams1, Nglobal = c(1, 4),
    type = "double")
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(streams): object 'streams1' not found}}\begin{verbatim}
as.vector(sim_1)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.vector(sim\_1): object 'sim\_1' not found}}\end{kframe}
\end{knitrout}
The arguments of the \fct{gpuRandom::runif} are described as follows:
\begin{itemize}
  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or specifying the length if to create a vector.
  \item \code{streams}: streams used for random number generation.
  \item \code{Nglobal}: global index space that defines the total work-items that execute in parallel. Default is set as $(64,8)$.
  \item \code{type}: ``double'' or ``float'' format of generated random numbers.
\end{itemize}
Next we show that by using a same stream \code{streams2} as \code{streams1},  \fct{gpuRandom::runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
sim_2 <- gpuRandom::runif(n = 6, streams = streams2, Nglobal = c(1, 4),
    type = "double")
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(streams): object 'streams2' not found}}\begin{verbatim}
as.vector(sim_2)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
\end{verbatim}
\end{kframe}
\end{knitrout}
Notice that \code{streams2} is a \code{deepcopy} of \code{streams1} before \code{streams1} is used. Reusing \code{streams1} will produce a vector different from \code{sim_1} (see \code{sim_3}). Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the first six rows (current states) of \code{streams1} after it being used.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
sim_3 = gpuRandom::runif(n = 6, streams = streams1, Nglobal = c(1, 4),
    type = "double")
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(streams): object 'streams1' not found}}\begin{verbatim}
as.vector(sim_3)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.vector(sim\_3): object 'sim\_3' not found}}\begin{verbatim}
t(as.matrix(streams1))[1:6, ]
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 't': object 'streams1' not found}}\end{kframe}
\end{knitrout}
Unlike the objects created by \proglang{R}, by default they remain in the memory after a restart. Streams created on GPU does not remain in the memory when a new \proglang{R} session begins. However, we can save the streams on the CPU in a file using the \fct{save} function, and later recall the streams with the \fct{load} function, in this way we can reproduce the same sequence of random numbers in simulations. Below is a trivial example, in which we demonstrate saving \code{streams1} to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU to produce the vector called \code{sim_saved}, which is also the same as \code{sim_1}.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
saveRDS(streams_from_GPU, "streams_from_GPU.rds")
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# Load the streams object as streams_saved
streams_saved <- gpuR::vclMatrix(readRDS("streams_from_GPU.rds"))
sim_saved = gpuRandom::runif(n=6,streams=streams_saved, 
                             Nglobal=c(1,4), type="double")
as.vector(sim_saved)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
\end{verbatim}
\end{kframe}
\end{knitrout}

We evaluate the distribution of 50000 GPU-generated double-precision uniform random numbers by making a density histogram  (the left panel in Figure \ref{fig0}), and plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel in Figure \ref{fig0}). The two cumulative distribution lines seem to overlap completely, suggesting the data generated by \fct{gpuRandom::runif} does come from a standard uniform distribution.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streams <- CreateStreamsGpu(Nstreams =512, keepInitial=1)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in CreateStreamsGpu(Nstreams = 512, keepInitial = 1): could not find function "{}CreateStreamsGpu"{}}}\begin{verbatim}
random_vector<-gpuRandom::runif(n=50000, streams=streams, 
                                Nglobal=c(64,8), type="double")
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(streams): object 'streams' not found}}\begin{verbatim}
random_vector<-as.vector(random_vector)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.vector(random\_vector): object 'random\_vector' not found}}\end{kframe}
\end{knitrout}
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in hist(random\_vector, prob = TRUE, main = "{}Histogram"{}): object 'random\_vector' not found}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in sort(x): object 'random\_vector' not found}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot.xy(xy.coords(x, y), type = type, ...): plot.new has not been called yet}}\end{kframe}
\end{knitrout}
\caption{histogram and empirical cumulative distribution of the GPU-generated random numbers.\label{fig0}}
\end{figure}

@@

In the following code, We create a $1024 \times 512$ matrix named \code{normal_matrix} of single-precision floating-point Gaussian random numbers, and assess Normality by making a histogram and a Q-Q plot of the generated random numbers in Figure \ref{fig2}. The points in the Q-Q plot shown in the right panel of Figure \ref{fig2} fall along the diagonal line, we have good evidence the generated data are from Gaussian distribution. We find doing a Q-Q plot can occupy a large amount of time in the process of producing an R markdown file, so as an incidental work, we adapted the \proglang{R} function \fct{stat::qnorm} to make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part on GPU, which can speed up the Q-Q plot greatly.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streams <- CreateStreamsGpu(Nstreams =512*128, keepInitial=1)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in CreateStreamsGpu(Nstreams = 512 * 128, keepInitial = 1): could not find function "{}CreateStreamsGpu"{}}}\begin{verbatim}
normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
                                Nglobal=c(512,128), type="float")
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(streams): object 'streams' not found}}\end{kframe}
\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
avector<-as.vector(as.matrix(normal_matrix))
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
hist(avector,breaks=40, main = "histogram")
gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
\end{verbatim}
\end{kframe}
\includegraphics[width=0.45\textwidth]{figure/histAndQQ-1} 
\includegraphics[width=0.45\textwidth]{figure/histAndQQ-2} 
\end{knitrout}
\caption{Histogram and Q-Q plot of the GPU-generated random numbers.\label{fig2}}
\end{figure}

@@

\subsection{Comparing running time: Month data example}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in CreateStreamsGpu(Nstreams = 128\textasciicircum{}2, keepInitial = 1): could not find function "{}CreateStreamsGpu"{}}}\end{kframe}
\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]
